---
title: 'Worksheet #3a'
author: "John Rick Estoce"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. You need to produce a vector that contains the first 11 letters.
```{r}
LETTERS[1:11]
```

b. Produce a vector that contains the odd numbered letters.
```{r}
LETTERS[seq(1, 26, 2)]
```

c. Produce a vector that contains the vowels
```{r}
LETTERS[LETTERS %in% c("A", "E", "I", "O", "U")]
```

d. Vector that contains the last 5 lowercase letters
```{r}
LETTERS[22:26]
```

e. Vector that contains letters between 15 to 24 (inclusive) in lowercase
```{r}
LETTERS[15:24]
```

##2. Create a vector(not a dataframe) with the average temperatures in April for Tugue-
garao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City. The average
temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.

a. What is the R code and its result for creating a character vector for the city/town
of Tuguegarao City, Manila, Iloilo City, Tacloban,Samal Island, and Davao City? Name the
object as city. The names should follow the same order as in the instruction.

```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban","Samal Island", "Davao City")
city
```

b. The average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.
Name the object as temp. Write the R code and its output. Numbers should also follow
what is in the instruction.

```{r}
temp <- c( 42, 39, 34, 34, 30, 27)
temp
```

c. Create a dataframe to combine the city and the temp by using ‘data.frame(). What
the R code and its result?
```{r}
data.frame(
  city = c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City"),
  temp = c(42, 39, 34, 34, 30, 27)
)
```


d. Associate the dataframe you have created in 2.(c) by naming the columns using
the names() function. Change the column names by using names() function as City and
Temperature. What is the R code and its result?
```{r}
df <- data.frame(
  city = c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City"),
  temp = c(42, 39, 34, 34, 30, 27)
)

names(df) <- c("City", "Temperature")

df
```


e. Print the structure by using str() function. Describe the output.
```{r}
str(df)
```
Explanation:
'data.frame': 6 obs. of 2 variables:
Your object df is a data frame with 6 observations (rows) and 2 variables (columns).

$ City : chr ...
The first column named City is of type character vector (chr).
It contains the city names as strings (like "Tuguegarao City", "Manila", etc.).

$ Temperature : num ...
The second column named Temperature is of type numeric (num).
It contains the average temperature values (42, 39, 34, etc.).

f. From the answer in d, what is the content of row 3 and row 4 What is its R code and its output?
```{r}
df[3:4, ]
```


g. From the answer in d, display the city with highest temperature and the city with the lowest temperature. What is its R code and its output?
```{r}
# City with highest temperature
df[df$Temperature == max(df$Temperature), "City"]

# City with lowest temperature
df[df$Temperature == min(df$Temperature), "City"]
```


##2. Create a matrix of one to eight and eleven to fourteen with four columns and three rows.
a. What will be the R code for the #2 question and its result?
```{r}
matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
```

b. Multiply the matrix by two. What is its R code and its result?
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

mat * 2
```
    

c. What is the content of row 2? What is its R code?
``` {r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)
mat[2, ]

```


d. What will be the R code if you want to display the column 3 and column 4 in row 1
and row 2? What is its output?
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

mat[1:2, 3:4]

```

e. What is the R code is you want to display only the columns in 2 and 3, row 3? What is its output?
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

mat[3, 2:3]
```


f. What is the R code is you want to display only the columns 4? What is its output?
``` {r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

mat[, 4]

```

g. Name the rows as isa, dalawa, tatlo and columns as uno, dos, tres, quatro for the matrix that was created in b.‘. What is its R code and corresponding output?
```{r}
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

rownames(mat) <- c("isa", "dalawa", "tatlo")
colnames(mat) <- c("uno", "dos", "tres", "quatro")

mat
```


h. From the original matrix you have created in a, reshape the matrix by assigning a new dimension with dim(). New dimensions should have 2 columns and 6 rows. What will be the R code and its output?
```{r}
# Original matrix
mat <- matrix(c(1:8, 11:14), nrow = 3, ncol = 4)

# Change dimensions to 6 rows and 2 columns
dim(mat) <- c(6, 2)

# Print the reshaped matrix
print(mat)
```


##3 An array contains 1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1

a. Create an array for the above numeric values. Each values will be repeated twice
What will be the R code if you are to create a three-dimensional array with 4 columns and
2 rows. What will be its output?

```{r}
# Original numeric values
values <- c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1)

# Repeat each value twice
values_repeated <- rep(values, 2)

# Create a 3-dimensional array with 2 rows, 4 columns, and 3 layers (because 24 / (2*4) = 3)
arr <- array(values_repeated, dim = c(2, 4, 3))

# Display the array
arr
```

b. How many dimensions do your array have?
The array you created has 3 dimensions.
```{r}
length(dim(arr))

```

c. Name the rows as lowercase letters and columns as uppercase letters starting from
the A. The array names should be “1st-Dimensional Array”, “2nd-Dimensional Array”, and
“3rd-Dimensional Array”. What will be the R codes and its output

```{r}
# Create the array (assuming previous code)
values <- c(1, 2, 3, 6, 7, 8, 9, 0, 3, 4, 5, 1)
values_repeated <- rep(values, 2)
arr <- array(values_repeated, dim = c(2, 4, 3))

# Name the rows, columns, and 3rd dimension
rownames(arr) <- letters[1:2]  # "a", "b"
colnames(arr) <- LETTERS[1:4]  # "A", "B", "C", "D"
dimnames(arr)[[3]] <- c("1st-Dimensional Array", "2nd-Dimensional Array", "3rd-Dimensional Array")

# Print the array
arr
```
