---
title: 'Worksheet #3a'
author: "John Rick Estoce"
date: "2025-10-13"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

a. You need to produce a vector that contains the first 11 letters.
```{r}
LETTERS[1:11]
```

b. Produce a vector that contains the odd numbered letters.
```{r}
LETTERS[seq(1, 26, 2)]
```

c. Produce a vector that contains the vowels
```{r}
LETTERS[LETTERS %in% c("A", "E", "I", "O", "U")]
```

d. Vector that contains the last 5 lowercase letters
```{r}
LETTERS[22:26]
```

e. Vector that contains letters between 15 to 24 (inclusive) in lowercase
```{r}
LETTERS[15:24]
```

2. Create a vector(not a dataframe) with the average temperatures in April for Tugue-
garao City, Manila, Iloilo City, Tacloban, Samal Island, and Davao City. The average

temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.

a. What is the R code and its result for creating a character vector for the city/town
of Tuguegarao City, Manila, Iloilo City, Tacloban,Samal Island, and Davao City? Name the
object as city. The names should follow the same order as in the instruction.

```{r}
city <- c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban","Samal Island", "Davao City")
city
```

b. The average temperatures in Celcius are 42, 39, 34, 34, 30, and 27 degrees.
Name the object as temp. Write the R code and its output. Numbers should also follow
what is in the instruction.

```{r}
temp <- c( 42, 39, 34, 34, 30, 27)
temp
```

c. Create a dataframe to combine the city and the temp by using ‘data.frame(). What
the R code and its result?
```{r}
data.frame(
  city = c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City"),
  temp = c(42, 39, 34, 34, 30, 27)
)
```

d. Associate the dataframe you have created in 2.(c) by naming the columns using
the names() function. Change the column names by using names() function as City and
Temperature. What is the R code and its result?
```{r}
df <- data.frame(
  city = c("Tuguegarao City", "Manila", "Iloilo City", "Tacloban", "Samal Island", "Davao City"),
  temp = c(42, 39, 34, 34, 30, 27)
)

names(df) <- c("City", "Temperature")

df
```
