---
title: 'WorkSheets#2'
author: "John Rick Estoce"
date: "2025-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Worksheet#2
##1. Create a vector using : operator
a. Sequence from -5 to 5. Write the R code and its output.
```{r setup, include=FALSE}
vec <- c(-5:5)
vec
```
output
[1] -5 -4 -3 -2 -1  0  1  2  3  4  5

```{r setup, include=FALSE}
 x <- c(1:7)
x
```
The value of x is 
[1] 1 2 3 4 5 6 7

##2  Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
Write the R code and its output.
```{r setup, include=FALSE}
sequence <- seq(1, 3, by = 0.2)
sequence
```
Each value is evenly spaced by 0.2.

##3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

```{r setup, include=FALSE}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
          22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43,
          53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48,
          27, 39, 19, 30, 61, 54, 58, 26, 18)
ages[3]
ages[c(2,4)]
ages[2:50]
```
output
a. [1] 22
b. [1] 28 36
c. [1] [1] 28 22 36 27 18 52 39 42 29 35 31 27 22 37 34 19 20 57 49 50 37 46 25 17 37 43 53 41 51 35 24 33 41
[34] 53 40 18 44 38 41 48 27 39 19 30 61 54 58 26 18

##4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
a. Print the results. Then access x[c("first", "third")].
Describe the output.

b. Write the code and its output.

```{r setup, include=FALSE}
sequence <- seq(1, 3, by = 0.2)
sequence
```
