---
title: 'WorkSheets#2'
author: "John Rick Estoce"
date: "2025-09-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Worksheet#2
##1. Create a vector using : operator
a. Sequence from -5 to 5. Write the R code and its output.
```{r setup, include=FALSE}
vec <- c(-5:5)
vec
```
output
[1] -5 -4 -3 -2 -1  0  1  2  3  4  5

```{r setup, include=FALSE}
 x <- c(1:7)
x
```
The value of x is 
[1] 1 2 3 4 5 6 7

##2  Create a vector using seq() function
a. seq(1, 3, by=0.2) # specify step size
Write the R code and its output.
```{r setup, include=FALSE}
sequence <- seq(1, 3, by = 0.2)
sequence
```
Each value is evenly spaced by 0.2.
Output: 
1.0  1.2  1.4  1.6  1.8  2.0  2.2  2.4  2.6  2.8  3.0

##3. A factory has a census of its workers. There are 50 workers in total. The following
list shows their ages: 34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35,
24,33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26,
18.

```{r setup, include=FALSE}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27,
          22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43,
          53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48,
          27, 39, 19, 30, 61, 54, 58, 26, 18)
ages[3]
ages[c(2,4)]
ages[2:50]
```
output
a. [1] 22
b. [1] 28 36
c. [1] [1] 28 22 36 27 18 52 39 42 29 35 31 27 22 37 34 19 20 57 49 50 37 46 25 17 37 43 53 41 51 35 24 33 41
[34] 53 40 18 44 38 41 48 27 39 19 30 61 54 58 26 18

##4. *Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the vector, names(x).
a. Print the results. Then access x[c("first", "third")].
Describe the output.

```{r setup, include=FALSE}
x <- c("first"=3, "second"=0, "third"=9)

print(x)

x[c("first", "third")]
```
b. Write the code and its output.
 first second  third 
     3      0      9 

 first third 
     3     9 
     
##5. Create a sequence x from -3:2.
a. Modify 2nd element and change it to 0;
b. Write the code and its output.
```{r setup, include=FALSE}
x <- c(-3:2)
x[2] <- 0
x
```
1st output[1] -3 -2 -1  0  1  2
2nd output: [1] -3  0 -1  0  1  2
The second element changed to zero after initializing the second element.

##6. *The following data shows the diesel fuel purchased by Mr. Cruz.
Month Jan Feb March Apr May June
Price per liter (PhP) 52.50 57.25 60.00 65.00 74.25 54.00
Purchase–quantity(Liters) 25 30 40 50 10 45

```{r setup, include=FALSE}
#a  Create a data frame for month, price per liter (php) and purchase-quantity (liter).
month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun")
price <- c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
liter <- c(25, 30, 40, 50, 10, 45)

fuel_data <- data.frame(month, price, liter)

fuel_data

#b. What is the average fuel expenditure of Mr. Cruz from Jan to June? Note: Use
#weighted.mean(liter, purchase)
avg_expenditure <- weighted.mean(price, liter)
avg_expenditure

```
Output:
a.
  month price liter
1   Jan 52.50    25
2   Feb 57.25    30
3   Mar 60.00    40
4   Apr 65.00    50
5   May 74.25    10
6   Jun 54.00    45

b.
[1] 59.2625

#7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths (in miles) of 141 “major” rivers in North America, as compiled by the US Geological Survey”.
a. Type “rivers” in your R console. Create a vector data with 7
elements, containing the number of elements (length) in rivers,#their sum (sum), mean (mean), median (median), variance (var)
standard deviation (sd), minimum (min) and maximum (max).
b. what's the output?
c. Code and its output
```{r setup, include=FALSE}

data <- c(
  length(rivers),
  sum(rivers),
  mean(rivers),
  median(rivers),
  var(rivers),
  sd(rivers),
  min(rivers),
  max(rivers)
)
data

#b. What's the output?
 # length   sum        mean  median  variance     sd       min     max 
 # 141.0   591952.0   591.2   425.0  98382.96   313.83   135.0  3710.0 


data <- c(
  length = length(rivers),
  sum = sum(rivers),
  mean = mean(rivers),
  median = median(rivers),
  variance = var(rivers),
  sd = sd(rivers),
  min = min(rivers),
  max = max(rivers)
)

data
 # length   sum        mean  median  variance     sd       min     max 
 # 141.0   591952.0   591.2   425.0  98382.96   313.83   135.0  3710.0 

```
##8. The table below gives the 25 most powerful celebrities and their annual pay as ranked by the editions of Forbes magazine and as listed on the Forbes.com website.

```{r setup, include= FALSE}
#a

power_rank <- 1:25

celebrity <- c(
  "Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods",
  "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos",
  "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali",
  "Paul McCartney", "George Lucas", "Elton John", "David Letterman",
  "Phil Mickelson", "J.K. Rowling", "Brad Pitt", "Peter Jackson",
  "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant" )

pay <- c(
  67, 90, 225, 110, 90,
  332, 302, 41, 52, 88,
  55, 44, 55, 40, 233,
  34, 40, 47, 75, 25,
  39, 45, 32, 40, 31
)
forbes <- data.frame(power_rank, celebrity, pay)
forbes

#b

jk_index <- which(forbes$celebrity == "J.K. Rowling")
forbes$power_rank[jk_index] <- 15
forbes$pay[jk_index] <- 90

forbes[jk_index, ]

#c.Interpret the data
#The outpot shows celebrities ranked by power and pay. After updating, J.K. Rowling is ranked 15th with a pay of 90 million dollars, putting her among other top earners. This shows how changes in pay can affect overall ranking.

```
