---
title: "RWorksheet_ESTOCE#4b.Rmd"
author: "John Rick Estoce"
date: "2025-11-03"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.
Hint Use abs() function to get the absolute
```{r}
# Define the vector
vectorA <- 1:5
matrix_result <- matrix(0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrix_result[i, j] <- abs(i - j)
  }
}

print(matrix_result)

```

2. Print the string "*" using for() function. The output should be the same as shown in Figure 2.
```{r}
for (i in 1:5) {
  for (j in 1:i) {
    cat('"*"\t')
  }
  cat("\n\n")
}

```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.
```{r}
fib_sequence <- function() {
  start <- as.integer(readline(prompt = "Enter starting number: "))
  
  if (start <= 0) {
    print("Please enter a positive number")
    return()
  }
  
  a <- 0
  b <- 1
  count <- 0
  result <- c()
  
  repeat {
    if (count >= start) {
      if (b > 500) {
        break
      }
      result <- c(result, b)
    }
    
    temp <- a + b
    a <- b
    b <- temp
    count <- count + 1
  }
  print(result)
}

```


4. Import the dataset as shown in Figure 1 you have created previously.

```{r}
library(readxl)
shoe_data <- read_excel("shoe_sizes.xlsx")
head(shoe_data)
```
